---
title: "Assignment 2"
author: "Haipeng Lin 44451253"
date: "05/09/2019"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
### Life A
Firstly, we need to find the birthday nearest the policy issue date as the integral insuring age.
The policy issue date was 20/06/1992, and the date of birth was 17/03/1964.
```{r}
as.Date("1992-06-20")-as.Date("1992-03-17")
```
```{r}
as.Date("1993-03-17")-as.Date("1992-06-20")
```
```{r}
1992-1964
```
The nearest birthday was 17/03/1992, and the insuring age on policy issue date was 28.
The observation period started on 20/06/1993. The initial exposed to risk of the survival at the insuring age of 30 was between 20/06/1994 and 20/06/1995. 
Therefore, the life A contributed 1 year to $E_{30}$.
```{r}
E30<-0
E30<-E30+1
E30
```

### Life B
Firstly, we need to find the birthday nearest the policy issue date as the integral insuring age.
The policy issue date was 06/08/1992, and the date of birth was 06/05/1964.
```{r}
as.Date("1992-12-18")-as.Date("1992-05-06")
```
```{r}
as.Date("1993-05-06")-as.Date("1992-08-06")
```
```{r}
1992-1964
```
The nearest birthday was 06/05/1992, and the insuring age on policy issue date was 28.
Life B dead on 12/06/1993 before the beginning of the observation period that was the policy anniversary in 1993 (06/08/1993). 
Therefore, the initial exposed to risk of the death at the insuring age of 30 should be 0, and the death should not be included in the number of death at insuring age of 30, $d_{30}$
```{r}
d30<-0
```

### Life C
Firstly, we need to find the birthday nearest the policy issue date as the integral insuring age.
The policy issue date was 18/12/1992, and the date of birth was 12/08/1964.
```{r}
as.Date("1992-12-18")-as.Date("1992-08-12")
```
```{r}
as.Date("1993-08-12")-as.Date("1992-12-18")
```
```{r}
1992-1964
```
The nearest birthday was 12/08/1992, and the insuring age on policy issue date was 28.
The observation period started on 18/12/1993. The survival actually exited on 18/06/1995. According to the definition given, the life C should exit on the policy anniversary in 1995 (18/12/1995) when the insuring age turned to 31. The initial exposed to risk of the survival at the insuring age of 30 is between 18/12/1994 and 18/12/1995. 
Therefore, the life C contributes 1 year to $E_{30}$.
```{r}
E30<-E30+1
E30
```

### Life D
Firstly, we need to find the birthday nearest the policy issue date as the integral insuring age.
The policy issue date was 04/01/1993, and the date of birth was 27/10/1964.
```{r}
as.Date("1993-01-04")-as.Date("1992-10-27")
```
```{r}
as.Date("1993-10-27")-as.Date("1993-01-04")
```
```{r}
1992-1964
```
The nearest birthday was 27/10/1992, and the insuring age on policy issue date was 28.
The observation period started on 04/01/1993. The initial exposed to risk of the survival at the insuring age of 30 was between 04/01/1995 and 04/01/1996. 
Therefore, the life D contributed 1 year to $E_{30}$.
```{r}
E30<-E30+1
E30
```

### Life E
Firstly, we need to find the birthday nearest the policy issue date as the integral insuring age.
The policy issue date was 28/04/1993, and the date of birth was 04/01/1965.
```{r}
as.Date("1993-04-28")-as.Date("1993-01-04")
```
```{r}
as.Date("1994-01-04")-as.Date("1993-04-28")
```
```{r}
1993-1965
```
The nearest birthday was 04/01/1993, and the insuring age on policy issue date was 28.
Life E dead on 29/08/1996 whose insuring age was 31, which should not be included in the number of death at insuring age of 30, $d_{30}$
The initial exposed to risk of the death at the insuring age of 30 was between 28/04/1995 and 28/04/1996.
Therefore, the life E contributed 1 year to $E_{30}$.
```{r}
E30<-E30+1
E30
```

### Life F
Firstly, we need to find the birthday nearest the policy issue date as the integral insuring age.
The policy issue date was 16/06/1993, and the date of birth was 18/04/1965.
```{r}
as.Date("1993-06-16")-as.Date("1993-04-18")
```
```{r}
as.Date("1994-04-18")-as.Date("1993-12-18")
```
```{r}
1993-1965
```
The nearest birthday was 18/04/1993, and the insuring age on policy issue date was 28.
The observation period started on 16/06/1993. The survival actually exited on 12/12/1995. According to the definition given, the life F should exit on the policy anniversary in 1995 (12/12/1995) when the insuring age turned to 30.Therefore, the initial exposed to risk of the survival at the insuring age of 30 is 0.

### Life G
Firstly, we need to find the birthday nearest the policy issue date as the integral insuring age.
The policy issue date was 29/10/1993, and the date of birth was 20/05/1965.
```{r}
as.Date("1993-10-29")-as.Date("1993-05-20")
```
```{r}
as.Date("1994-05-20")-as.Date("1993-10-29")
```
```{r}
1993-1965
```
The nearest birthday was 20/05/1993, and the insuring age on policy issue date was 28.
The initial exposed to risk of the death at the insuring age of 30 was between 29/10/1995 and 29/10/1996.
Life G dead on 21/04/1996 whose insuring age was 30, which should  be included in the number of death at insuring age of 30, $d_{30}$
Therefore, the life G contributed 1 year to $E_{30}$.
```{r}
E30<-E30+1
E30
```
```{r}
d30<-d30+1
d30
```

### Life H
Firstly, we need to find the birthday nearest the policy issue date as the integral insuring age.
The policy issue date was 16/02/1994, and the date of birth was 04/07/1965.
```{r}
as.Date("1994-02-16")-as.Date("1993-07-04")
```
```{r}
as.Date("1994-07-04")-as.Date("1994-02-16")
```
```{r}
1994-1965
```
The nearest birthday was 04/07/1994, and the insuring age on policy issue date was 29.
 The initial exposed to risk of the survival at the insuring age of 30 was between 16/02/1995 and 16/02/1996. 
Therefore, the life H contributed 1 year to $E_{30}$.
```{r}
E30<-E30+1
E30
```
### Life I
The life I contributed 1 year to $E_{30}$.
```{r}
E30<-E30+1
E30
```
### Life J
The death be included in the number of death at insuring age of 30, $d_{30}$
```{r}
d30<-d30+1
d30
```
```{r}
E30<-E30+1
E30
```
The mortality is $q_{30}=\frac{d_{30}}{E_{30}}$
```{r}
q30<-d30/E30
q30
```

## Question 2

The initial exposed to risk, actual deaths and expected deaths of the grouping data will be imported into the software.

Age group
```{r}
x<-c("40–44","45–49","50–54","55–59","60–64","65–69","70–74","75–79","80–84","85–89","90–94")
x
```
Exposed to risk (years) $E_x$
```{r}
Ex<-c(15518,19428,21594,21890,19174,15775,11414,6993,3276,1096,201)
Ex
```
Actual deaths $dx$
```{r}
dx<-c(65,144,219,378,465,557,685,644,471,217,67)
dx
```
Expected deaths $Edx$
```{r}
Edx<-c(73.9,134.6,223.9,346.3,468.1,600.2,675.5,637.4,458.7,240.6,61.4)
Edx
```
Graduated $\dot{q_x}$
```{r}
qx<-Edx/Ex
qx
```
The variance at each age of deaths $E_x\dot{q_x}(1-\dot{q_x})$
```{r}
VDeaths<-Edx*(1-qx)
VDeaths
```
The standardised deviation$$Z_x=\frac{d_x-E_x\dot{q_x}}{\sqrt{E_x\dot{q_x}(1-\dot{q_x})}}$$
```{r}
Zx<-(dx-Edx)/sqrt(Edx*(1-qx))
Zx
```

### Statistical Test
After importing the above data and obtaining the required values, we can start the statistical test.

The null hypothesis for all the tests is that the graduation adheres closely to the data.

### Chi Square Test
The test statistic is:$$\chi^2=\sum_{all\,ages}Z_x$$
```{r}
chi_2<-sum(Zx^2)
chi_2
```
Under the null hypothesis, this test statistic has a $\chi^2$ distribution with 11 degrees of freedom at the 5% significance level, i.e. $\chi^2_{0.95,11}$.
```{r}
qchisq(0.95,11)
```
Since the vealue of the test statistic $\chi^2$ is smaller than 19.67, we cannot reject the null hypothesis at the 5% level of significance and conclude that the graduation adheres closely to the data. 
The chi square test cannot detect some large deviations balanced by lots of smaller ones.

### Indibidual Standardised Deviations Test
The test statistic is:$$\sum_{all\,intervals}\frac{(A-E)^2}{E}$$ $A$ is the actual number and $E$ is the expected number.

Firstly, we need the probabilities that the $Z$-scores are expected to be in various ranges. Since the range of the $Z$-scores we have is (-2,2) that only 4 ranges are needed, i.e.  
($-\infty$,-1), (-1,0), (0,1), (1,$\infty$).
```{r}
r<-c()
r[1]<-pnorm(-1)
r[2]<-pnorm(0)-pnorm(-1)
r[3]<-pnorm(1)-pnorm(0)
r[4]<-1-pnorm(1)
r
```
The expected distribution of the standardised deviation is as follow.
```{r}
Er<-11*r
Er
```
The actual distribution of the standardised deviation is as follow.
```{r}
Ar<-c()
Ar[1]<-sum(Zx<(-1))
Ar[2]<-sum((Zx>-1)*Zx<(0))
Ar[3]<-sum((Zx>0)*(Zx<(1)))
Ar[4]<-sum(Zx>1)
Ar
```
Overall,there are 5 negatives and 6 positives, and the actual distribution looks a little different compared to the expected, hence there is weak evidence of non-adherence to data. 
```{r}
sum((Zx>(-2/3))*(Zx<(2/3)))
```
There are 5 deviations in the range $(-\frac{2}{3},\frac{2}{3})$. We would expect to have 50% of the 11 deviation (i.e. 5.5) in this range that there is no significant difference.
```{r}
ISD<-sum((Ar-Er)^2/Er)
ISD
```
The value of the test statistic should be compared with the critical value whchi is \chi^2 with 3 degrees of freedom at 5% significance level, i.e. $\chi^2_{0.95,3}$.
```{r}
qchisq(0.95,3)
```
Since the critical value is greater than the test statistic value, we cannot reject the null hypothesis and conclude that the graduation adherence to data is good. 
This test fails to detect the clumping of deviations of same sign because of the insufficient amount of data.

### Cumulative Deviation Test
The test statistic is:$$\frac{\sum_{x}(d_x-E_x\dot{q_x})}{\sqrt{\sum_{x}E_x\dot{q_x}(1-\dot{q_x})}}$$
```{r}
CDT<-sum(dx-Edx)/sqrt(sum(Edx*(1-qx)))
CDT
```
This method is two-tailed, and its observed test statistic is -0.14. We need to compare it with the critical value $\Phi(0.975)$ at 5% significance level in absolute value.
```{r}
CDT<-abs(CDT)
CDT
```
```{r}
qnorm(0.975)
```
Therefore, the observed test statistic is much smaller than the critical value, we cannot reject the null hypothesis at the 5% level of significance and it has a standard normal sampling distribution. 
This test will not detact clumping of deviation of the same sign and outliers if they lie either side of the curve.

### Sign Test
If the null hypothesis is true, we would expect that roughly half the crude rates would above the expected rates and half below. The number of positive standardised deviations will have a $Binomial(11,0.5)$.
```{r}
sum((Zx>0))
```
There are 6 positive standardised deviations.$$P(Observed\,number\,of\,positive\,deviations\leq 6)$$
```{r}
pbinom(6,11,0.5)
```
This is a two-side test. So the $p$-value is
```{r}
2*pbinom(6,11,0.5)
```
Since the  $p$-value is greater than 5%, there is no reason to reject the null hypothesis.

### Grouping of Signs Test \frac{1}{k}
If the null hypothesis is true, we would expect that the number of distinct groups of positive $Z_x$ should be larger than k, which is smallest possible value for$$\sum_{j=1}^{k}\frac{\begin{pmatrix}n_1-1\\\\j-1\end{pmatrix}\begin{pmatrix}n_2+1\\\\j\end{pmatrix}}{\begin{pmatrix}n_1+n_2\\\\n_1\end{pmatrix}}\geq0.05$$n_1 is number of positive deviations, n_2 is number of negative deviations
```{r}
Zx
```
There are 4 groups of standardised deviatiion, 6 positive standardised deviations, and 5 negatives.
```{r}
n1<-6;n2<-5
j<-0
repeat{j<-j+1
  k<-c(1:j)
if((sum(factorial(n1-1)/factorial(k-1)/factorial(n1-k)*factorial(n2+1)/factorial(k)/factorial(n2+1-k)))/(factorial(n1+n2)/factorial(n1)/factorial(n2))>=0.05){print(j)
  break}
}
```
The number of distinct groups of pictive $Z_x$ is larger than 2, so there is no reason to reject the null hypothesis.
This test fails to detect excessive absolute deviations.

### Serial Correlations Test
The test statistic is:$$\frac{\sum_{i=1}^{m-1}(z_i-\bar{z}^{(1)})(z_{i+1}-\bar{z}^{(2)})}{\sqrt{\sum_{i=1}^{m-1}(z_i-\bar{z}^{(1)})^2\sum_{i=1}^{m-1}(z_{i+1}-\bar{z}^{(2)})^2}}\sqrt{m-1}$$
$$\bar{z}^{(1)}=\frac{\sum_{i=1}^{m-1}z_i}{m-1}\  \bar{z}^{(2)}=\frac{\sum_{i=1}^{m-1}z_{i+1}}{m-1}$$
```{r}
zbar_1=Zx[1:10]
zbar_2=Zx[2:11]
SCT<-sum((zbar_1-mean(zbar_1))*(zbar_2-mean(zbar_2)))/sqrt(sum((zbar_1-mean(zbar_1))^2)*sum((zbar_2-mean(zbar_2))^2))*sqrt(10)
SCT
```
The value of test statistic should be compare with $\Phi(0.95)$.

```{r}
qnorm(0.95)
```
Since the test statistic value is much smaller than the critical value, so there is no reason to reject the null hypothesis. 

### Conclusion
Overall, we do not have sufficient evidence to reject our null hypothesis, and the graduation adheres closely to the data. 

## Question 3
### Tutorial 5.7
Firstly, we import the data needed.
Age $x$
```{r}
x<-c(70:75)
x
```
Central Exposed to Risk $E^C_x$
```{r}
ECx<-c(1000,1005,1010,1008,1006,998)
ECx
```
Observed Number of Death $d_x$
```{r}
dx<-c(80,90,95,105,115,125)
dx
```
Crude $\mu_x$
```{r}
mu_x<-dx/ECx
mu_x
```
In order to obtain the graduated estimates of force of mortality $\dot{\mu}$ for the above data, we need to use the poisson model to find its maximum likelihood estimate. Since the age label of the data is $\prime$aged $x$ nearest birthday$\prime$, the $\dot\mu$-estimate applies to age $x$.  $$P_r(D_x=d_x)=\frac{e^{-E^C_x\mu_x}(E^C_x\mu_x)^{d_x}}{d_x!}$$
Its log likelihood functioin is $$l(\mu)=-\sum_{x}E^C_x\mu_x+\sum_{x}d_xln(E^C_x\mu_x)$$
According to Gompertz$\prime$ Law, $\mu_x=Bc^x$.
$$l(\mu)=-\sum_{x}E^C_xBc^x+\sum_{x}d_xln(E^C_xBc^x)$$
Then $$\frac{\partial{l(\mu)}}{\partial{B}}=-\sum_{x}E^C_xc^x+\sum_{x}\frac{d_x}{B}$$
$$\frac{\partial{l(\mu)}}{\partial{c}}=-\sum_{x}E^C_xBxc^{x-1}+\sum_{x}d_x\frac{x}{c}$$
Let these two equations be 0, and then we can get the MLE of parameter $B$ and $c$, which can be calculated by the R function for non-linear minimisation function, $nlm()$.
In order to achieve better estimates of $B$ and $c$, we can construct a linear regression model to find the initial starting point.$$ln(\hat{\mu_x})=lnB+xlnc$$
We can use the crude $\hat{\mu_x}$ to help construct the linear regression model.
```{r}
y<-log(mu_x)
y
```

```{r}
mod_1<-lm(y~x)
mod_1
```
Therefore, $lnB$ and $lnc$ are
```{r}
lnB<-mod_1$coefficients[1]
lnB
```
```{r}
lnc<-mod_1$coefficients[2]
lnc
```
The initial points of $B$ and $c$ are
```{r}
B_in<-exp(lnB)
B_in
```
```{r}
c_in<-exp(lnc)
c_in
```
We can use these two initial values and $nlm()$ functioin to find the exact value of $B$ and $c$.
```{r}
f_1<-function(p){abs(sum(-ECx*p[1]*p[2]^x+dx*log(ECx*p[1]*p[2]^x,base=exp(1))))}
iv<-c(B_in,c_in)
para<-nlm(f_1,iv)
para
```
The MLE of $B$ and $c$ are
```{r}
B_e<-para$estimate[1]
B_e
```

```{r}
c_e<-para$estimate[2]
c_e
```
The graduated $\mu_x$ are
```{r}
mu_grad<-B_e*c_e^x
mu_grad
```

### Tutorial 5.8
Firstly, we import the data needed.
Age $x$
```{r}
x<-c(30:49)
x
```
Initial Exposed to Risk $E_x$
```{r}
Ex<-c(70000,66672,68375,65420,61779,66091,68514,69560,65000,66279,67300,65368,65391,62917,66537,62302,62145,63856,61097,61110)
Ex
```
Observed Number of Death $d_x$
```{r}
dx<-c(39,43,34,31,23,50,48,43,48,47,62,63,84,86,120,121,122,162,151,184)
dx
```
Crude $q_x$
```{r}
q_x<-dx/Ex
q_x
```
In order to obtain the graduated estimates of mortality $\dot{q}$ for the above data, we use the model provided and weighted least squares. Since the age label of the data is $\prime$aged $x$ last birthday$\prime$, the $\dot\q$-estimate applies to age $x$. The weighted least squares is $$\sum_xW_x(\hat{q_x}-\dot{q_x})^2$$
$$W_x=E_x$$.

The provided model is $$ln(\frac{\dot{q_x}}{1-\dot{q_X}})=a+bx$$.

Therefore, in order to achieve the better graduated estimates, we can use crude $\hat{q_x}$ and the above model to find the initial values of $a$ and $b$ for $nlm()$ function.
$$ln(\frac{\hat{q_x}}{1-\hat{q_X}})=a+bx$$

```{r}
y<-log(q_x/(1-q_x))
y
```

```{r}
mod_2<-lm(y~x)
mod_2
```
Therefore, $a$ and $b$ are
```{r}
a_in<-mod_2$coefficients[1]
a_in
```
```{r}
b_in<-mod_2$coefficients[2]
b_in
```
We can use these two initial values and $nlm()$ functioin to find the exact value of $a$ and $b$.
```{r}
f_2<-function(p){abs(sum(Ex*(log(q_x/(1-q_x))-(p[1]+p[2]*x))^2))}
iv<-c(a_in,b_in)
para_2<-nlm(f_2,iv)
para_2
```
The MLE of $a$ and $b$ are
```{r}
a_e<-para_2$estimate[1]
a_e
```

```{r}
b_e<-para_2$estimate[2]
b_e
```
$$ln(\frac{\dot{q_x}}{1-\dot{q_X}})=\hat{a}+\hat{b}x$$
The graduated $\dot{q}_x$ are
```{r}
qx_grad<-1-1/(1+exp(a_e+b_e*x))
qx_grad
```

### Tutorial 5.9
Firstly, we import the data needed.
Age $x$
```{r}
x<-c(47:67)
x
```
Initial Exposed to Risk $E_x$
```{r}
Ex<-c(166,187,218,243,276,302,347,390,430,494,558,628,701,813,917,1040,1182,1299,1432,1596,1752)
Ex
```
Observed Number of Death $d_x$
```{r}
dx<-c(2,2,4,6,2,4,7,3,9,9,8,11,14,18,18,24,30,43,41,54,64)
dx
```
Standard Table Figure $q^S_x$
```{r}
qsx<-c(0.00505,0.00570,0.00644,0.00728,0.00826,0.00930,0.01051,0.01184,0.01331,0.01492,0.01668,0.01859,0.02065,0.02287,0.02525,0.02778,0.03049,0.03339,0.03648,0.03978,0.04332)
qsx
```
Crude $q_x$
```{r}
q_x<-dx/Ex
q_x
```
In order to obtain the graduated estimates of mortality $\dot{q}$ for the above data, we use the model provided and weighted least squares. Since the age label of the data is $\prime$aged $x$ last birthday$\prime$, the $\dot\q$-estimate applies to age $x$. The weighted least squares is $$\sum_xW_x(\hat{q_x}-\dot{q_x})^2$$
$$W_x=E_x$$.

The provided model is $$\dot{q_x}=a+bq^S_x$$.

Therefore, in order to achieve the better graduated estimates, we can use crude $\hat{q_x}$ and the above model to find the initial values of $a$ and $b$ for $nlm()$ function.
$$\hat{q_x}=a+bq^S_x$$

```{r}
y<-q_x
y
```

```{r}
mod_3<-lm(y~qsx)
mod_3
```
Therefore, $a$ and $b$ are
```{r}
a_in<-mod_3$coefficients[1]
a_in
```
```{r}
b_in<-mod_3$coefficients[2]
b_in
```
We can use these two initial values and $nlm()$ functioin to find the exact value of $a$ and $b$.
```{r}
f_2<-function(p){abs(sum(Ex*(q_x-(p[1]+p[2]*qsx))^2))}
iv<-c(a_in,b_in)
para_3<-nlm(f_2,iv)
para_3
```
The MLE of $a$ and $b$ are
```{r}
a_e<-para_3$estimate[1]
a_e
```

```{r}
b_e<-para_3$estimate[2]
b_e
```
$$\dot{q_x}=a+bq^S_x$$.
The graduated $\dot{q}_x$ are
```{r}
qx_grad<-a_e+b_e*qsx
qx_grad

```

